
---
title: "Chapter 4. The effect of verb class"
author: "Raquel Montero Estebaranz"
date: "05.22.2024"
output: 
  html_document: 
      toc: true
      toc_float: true
      theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center", warning = FALSE, message = FALSE)

```
```{css, echo=FALSE}
h1,h4 {
  text-align: center;
}
p {
  text-align: justify; 
}
```

## 1. Introduction

This file shows the code used to analyse the data for Chapter 4 of the thesis "Mood alternations: a synchronic and diachronic study of negated complement clauses". The data as well as the annotation guidelines can be found at: https://github.com/Raquel-Montero


The following are the Packages that will be used:
```{r}
library(readr)
library(carData)      # for cat package
library(car)          # Anova function
library(dplyr)        # Operations
library(plyr)         # for ddply 
library(ggplot2)      # to use ggplot
library(sjPlot)       # to change the font
library(Matrix)       # for lme4 package
library(lme4)         # to calculate lmer models
library(lmerTest)     # for the p values
library(mgcv)         # for Gam model
library(itsadug)      # for Gam model
library(tidygam)      # for Gam plotting
library(tidymv)       # For Gam plotting: https://www.rdocumentation.org/packages/tidymv/versions/3.4.2
library(ggpubr)
library(LaplacesDemon) #for the function invlogit
library(ggpubr)
library(functional)
library(scales)
```


## 2. General Development: Graphs

Loading the Data: 

```{r}
data <- read_csv("DiachronyAllVerbs.csv", 
                  show_col_types = FALSE,
                  locale = locale(encoding =  "ISO-8859-1"))

```

Final Data that will be used for analysis:

```{r}
clean.data <- subset(data, Mverbtype =="ind" & # the matrix verb is in indicative
                     Mverbclass!="Na"& # excludes any data that is not one of the four verbs
                     Emood !="Na" & # no embedded verb
                     Emood !="NA" &
                     Emood !="inf" & # embedded verb in infinitive
                     MClauseType=="noninterrogative" & # only takes into account non-interrogative clauses
                     EClauseType=="unambiguous" # exclude ambigous sentences
                   )

# Changing the indicative to 1 and subjunctive to 0:
clean.data$Emood2 <- ifelse(clean.data$Emood == "subj", 0, 1)

# Converting the mood into a numeric value:
clean.data$Emood2 <- as.numeric(as.character(clean.data$Emood2))

#changing names of values so that they are clearer: 
clean.data$Mverbclass2 <- ifelse(clean.data$Mverbclass == "factive", "semi-factive", "non-factive")
clean.data$Construction <- ifelse(clean.data$Construction == "1st present", "1st present (FP)", "other (O)")

```

Data for plotting: 

```{r}
# Creates new data-frame with the means of embedded mood per period, verb class/verb and construction:

plot.data.verbtype <- ddply(clean.data, # data frame
                          .(Period, Construction,Mverbclass2), # variables to calculate the mean
                          summarize, #summary
                          mean = mean(as.numeric(as.character(Emood2)), na.rm = T), # mean
                          n = sum(!is.na(as.numeric(as.character(Emood2)))) # number of tokens
                          )
plot.data.verbtype

plot.data.verbs <- ddply(clean.data, 
                        .(Period, Construction,Mverbl), 
                        summarize, 
                        mean = mean(as.numeric(as.character(Emood2)), na.rm = T),
                        n = sum(!is.na(as.numeric(as.character(Emood2))))
                        )
plot.data.verbs
```

Plotting the results:

```{r}
#Font and theme for the graph:
windowsFonts("Garamond" = windowsFont("Garamond"))

set_theme( base = theme_bw(base_family = 'Garamond'),
  title.size = 1.4,
  axis.title.size = 1.4,
  axis.textsize = 1.1,
  legend.size = 1,
  legend.title.size = 1,
  geom.label.size = 3
)

# Plot 1: Mood selection of different verb classes

plot.data.verbtype$Period <- as.factor(plot.data.verbtype$Period) # Period as factor

verbtype.diachrony <- ggplot()+  
                geom_point(data=plot.data.verbtype, #main data
                           aes(Period, # x-axis
                               mean, # y-axis
                               size = n, # size of the dots
                               color=Construction #colors
                               )
                           )+  
                facet_wrap(~factor(Mverbclass2, # divide into facets
                                   levels = c('semi-factive', 'non-factive') # order of the facets
                                   )
                           )+ 
                scale_size_area(max_size=13)+  # controls the maximum side of the points.
                labs(title="Proportion Mood: not V(ind) that p(ind/subj)",  # axis labels
                      x = " ", 
                       y="Mood Proportion (1=ind,0=subj)")+
                scale_color_manual(values=c("black", "#5F9EA0" ))+ # color values
                labs(size="size", colour="Construction")+ # labels legends
                ylim(0,1) # minimum/maximum values of the y-axis
  
verbtype.diachrony

# Plot 2: mood selection after each individual verb

plot.data.verbs$Period <- as.factor(plot.data.verbs$Period) # Period as factor

verbs.diachrony <- ggplot()+  
              geom_point(data=plot.data.verbs, #main data
                         aes(Period, # x-axis
                             mean, # y-axis
                             size = n, # size of the dots
                             color=Construction) # colors
                         )+
             facet_wrap(~factor(Mverbl, #divide into facets
                                levels=c('saber (know)', 'creer (believe)', 'ver (see)', 'decir (say)')), #order facets
                                ncol=2 # number of columns
                        )+ 
              scale_size_area(max_size=13)+ # maximum size of the dots
              labs(title="Proportion Mood: not V(ind) that p(ind/subj)", #labels 
                    x = " ", 
                    y="Mood Proportion (1=ind,0=subj)")+
             scale_color_manual(values=c("black", "#5F9EA0" ))+ # colors plot
            ylim(0,1)+ # minimum/maximum value of the y-axis
            labs(size="size", colour="Construction")
verbs.diachrony
```
```{r}
#Saving the plots:
ggsave(verbtype.diachrony, file="verb-classes-diachrony.png", width = 8, height= 4)
ggsave(verbs.diachrony, file="verbs-diachrony.png", width = 8, height= 6)
```



## 3. Statistical Analysis
```{r}
# changing verbclass and construction to factor:
clean.data$Mverbclass <- as.factor(clean.data$Mverbclass)#
clean.data$Construction <- as.factor(clean.data$Construction)

# change year into a numeric variable:
clean.data$Year2 <- as.numeric(clean.data$Year2)

# create a new column with Class and Construction as one combined factor: (for the GAM model)
clean.data <- mutate(clean.data, classContruction = interaction(Mverbclass, Construction))

# use a contrast treatment and order the variable:
clean.data$classContruction <- as.ordered(clean.data$classContruction)
contrasts(clean.data$classContruction) <- "contr.treatment"

```



### 3.1. GAM: Exploring the verb class effect


#### Gam models
```{r}
# No effect of Verb class:
model <- bam(Emood2 ~ Construction+
                     s(Year2)+
                    s(Year2,by=Construction),
                   method = "REML", 
                  family="binomial",
                   data = clean.data)

# Main efect of verb class but no interaction with it: --> 
model2 <- bam(Emood2 ~ Construction+Mverbclass+
                     s(Year2)+
                     s(Year2, by=Construction),
                   method = "REML", 
                  family="binomial",
                   data = clean.data)


# Three way interaction:
model3 <- bam(Emood2 ~ classContruction+
                     s(Year2)+
                     s(Year2, by=classContruction),
                   method = "REML", 
                  family="binomial",
                   data = clean.data)


```

#### Comparision of the models

```{r}
# Comparing the models: 
AIC(model2, model) # no effect of verb class vs. main effect of verb class
AIC(model3, model2) # main effect of verb class vs. three way interaction

```



#### Gam final model

```{r}

final.model <- bam(Emood2 ~ classContruction+
                     s(Year2)+
                     s(Year2, by=classContruction),
                   method = "REML", 
                  family="binomial",
                   data = clean.data)

summary(final.model)

```
Checking the model results: 
```{r}
gam.check(final.model)
k.check(final.model)
```



#### Ploting the GAM
```{r}

plot.binomial <- plot_smooths(model = final.model, 
             series = Year2,
             comparison = Construction, 
             facet_terms = Mverbclass2, 
             split = list(classContruction = c( "Mverbclass2","Construction")), 
             transform = invlogit # the model log transforms the data, to plot the odds ratio we use the inverse f
             ) +
  scale_color_manual(values = c("black", "#5F9EA0"), name = "Construction") + 
  scale_fill_manual(values = c("gray80", "#5F9EA0"), name = "Construction") +
  scale_linetype_manual(values = c("solid", "dotted"), name = "Construction") +
  theme_classic() + 
  ylab("Predicted Proportion Mood") + 
  xlab("Years")

plot.binomial

```
```{r}
#saving the plot:
ggsave(plot.binomial, file="plot-gam-binomial.png", width = 8, height = 4)
```


#### Difference curve

```{r}

# Plotting the difference curve: 

combined <- ggarrange(plot_difference(final.model, Year2, 
                          difference= list(classContruction = 
                                             c("factive.other (O)",
                                               "factive.1st present (FP)"))) + 
            scale_y_continuous(limits=c(-2, 7), breaks = seq(-1, 7, by = 1))+
            ylab("Difference FP and O") + 
            xlab("Years") + 
            theme_classic() + 
            ggtitle("Construction O minus Construction FP | semi-factive verbs"),
            plot_difference(final.model, Year2, 
                          list(classContruction = c("non-factive.other (O)",
                                                    "non-factive.1st present (FP)"))) + 
             scale_y_continuous(limits=c(-2, 7),breaks = seq(-1, 7, by = 1))+
            ylab("Difference FP and O") + 
            xlab("Years") + 
            theme_classic() + 
            ggtitle("Construction O minus Construction FP | non-factive verbs"))
combined
            
```

```{r}
ggsave(combined, file="combined.png", width = 11, height = 4)
```

 Plotting the Difference Curve with a change in the scale

```{r}
#Predictions of the model:
preds <- get_gam_predictions(final.model, 
                             Year2, 
                             split = 
                  list(classContruction = c("Verb Class",
                                            "Constrcution")))

# Backtransforming the values:
preds$Emood2<-invlogit(preds$Emood2)

# Change into a data-frame:
my_df <- as.data.frame(preds)

# diff: 2-1 by Year2
my_df1 <- subset(my_df, .idx=="1")
my_df2 <- subset(my_df, .idx=="2")
differenceV1 <- (my_df2$Emood2 - my_df1$Emood2)

# Diff: 4-3 by Year2
my_df3 <- subset(my_df, .idx=="3")
my_df4 <- subset(my_df, .idx=="4")
differenceV2 <- (my_df4$Emood2 - my_df3$Emood2)

# Years
Years <- my_df3$Year2

# V1 significance level: 1360 onwards significant
significanceV1 <- c(FALSE,FALSE,FALSE,FALSE,FALSE,
                    TRUE, TRUE, TRUE, TRUE, TRUE,
                    TRUE, TRUE, TRUE, TRUE, TRUE,
                    TRUE, TRUE, TRUE, TRUE, TRUE,
                    TRUE, TRUE, TRUE, TRUE, TRUE)

# V2 significance level: 
significanceV2 <- c(FALSE,FALSE,FALSE,
                    TRUE, TRUE, TRUE, TRUE,
                    TRUE, TRUE, TRUE, 
                    FALSE,FALSE,FALSE,
                    TRUE, TRUE, TRUE, TRUE, TRUE,
                    TRUE, TRUE, TRUE, TRUE, TRUE,
                    TRUE,TRUE)

# combining all the above
newdata <- cbind(Years,differenceV1,significanceV1,differenceV2,significanceV2)
newdata <- as.data.frame(newdata) #transform into data frame


#we can now plot these differences:
#Plot 1:
 PlotV1 <- ggplot(newdata, aes(Years, differenceV1))+
   geom_line()+
   geom_hline(aes(yintercept = 0), colour = "black") +
  geom_rect(aes(xmin = 1360, xmax = 2000, ymin = -Inf, ymax = Inf),
                   fill = "red", alpha = 0.005)+
   ylim(-0.2,1)+
   ylab("Difference: O-FP")+
   ggtitle("Construction O minus Construction FP|semi-factive verbs")+
   theme_bw() + 
   theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

 # Plot 2:
 PlotV2 <- ggplot(newdata, aes(Years, differenceV2))+
   geom_line()+
    geom_hline(aes(yintercept = 0), colour = "black") +
   geom_rect(aes(xmin = 1256, xmax = 1512, ymin = -Inf, ymax = Inf),
                   fill = "red", alpha = 0.005)+
   geom_rect(aes(xmin = 1625, xmax = 2000, ymin = -Inf, ymax = Inf),
                   fill = "red", alpha = 0.005)+
   ylim(-0.2,1)+
   ylab("Difference: O-FP")+
   ggtitle("Construction O minus Construction FP|non-factive verbs")+
   theme_bw() + 
   theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
 
 #combining the plots
 combinedraw <- ggarrange(PlotV1,PlotV2)
 combinedraw


```
```{r}
ggsave(combinedraw,file="combined-difference-curve.png", width = 11, height = 4)
```


### 3.2. GAM: exploring the effect of individual predicates


```{r}
# Create a new column with Class and Construction as one combined factors:
clean.data <- mutate(clean.data, PredicateConstruction = interaction(Mverbl, Construction))

#use a contrast treatment and order the variable:
clean.data$PredicateContruction  <- as.ordered(clean.data$PredicateConstruction )
contrasts(clean.data$PredicateConstruction ) <- "contr.treatment"

model.predicates <- bam(Emood2 ~ PredicateConstruction  +
                     s(Year2) +
                     s(Year2, by = PredicateConstruction),
                   method = "REML", 
                   family = "binomial",
                   data = clean.data)

#comparing the models: 
AIC(final.model,model.predicates)


```
The model in which individual predicates rather than verb class is significantly better. It must be said that the data is considerably less for each individual predicate. 
The work leaves open the possibility of exploring theoretical specific difference across verbs within the same verbal class.

## 4. Construction FP: semi-factive

Data for first person construction semi-factive verbs:
```{r}
# Subseting the data:

clean.data.factive <- subset(clean.data, 
                     Mverbclass=="factive"& # Only factive verbs
                     Construction=="1st present (FP)"
                   )

plot.data.verbs.factive <- ddply(clean.data.factive, .(Period,Mverbl), 
                        summarize, 
                        mean = mean(as.numeric(as.character(Emood2)), na.rm = T),
                        n = sum(!is.na(as.numeric(as.character(Emood2))))
                        )
# know:
clean.data.know <- subset(clean.data, 
                    Mverbl=="saber (know)"& # Only the verb know
                    Construction=="1st present (FP)"
                   )
# see: 
clean.data.see <- subset(clean.data, 
                    Mverbl=="ver (see)"& # Only the verb see
                    Construction=="1st present (FP)"
                   )

```
#### First attempt: fitting all data

```{r}
# Logarithm:
logarithm <- function(Period,k,s)(1/(1+exp(s*(Period-k))))

# model:
model <- nls(Emood2 ~ logarithm(Period,k,s), data=clean.data.know, start=list(s=0.024, k=1439))
summary(model) 

#values of the model: 0.019+-0.003 [0.22-0.016], k=1428+-12 [1440-] 

modelsee <- nls(Emood2 ~ logarithm(Period,k,s), data=clean.data.see, start=list(s=0.008, k=1505))
summary(modelsee) 

#values of the model: s=0.008+-0.002[0.01-0.006] k=1506.+-23

```
```{r}
clean.data.know$Period<-as.numeric(clean.data.know$Period)
plot.data.verbs.factive$Period <-as.numeric(plot.data.verbs.factive$Period)

# Functions for plotting
logsaber <-function(x) {1/(1+exp(0.019*(x-1428)))}
logsaberupper <-function(x) {1/(1+exp(0.025*(x-1452)))}
logsaberlower <-function(x) {1/(1+exp(0.013*(x-1404)))}


logver <-function(x) {1/(1+exp(0.008*(x-1506)))}
logverupper <-function(x) {1/(1+exp(0.012*(x-1552)))}
logverlower <-function(x) {1/(1+exp(0.004*(x-1460)))}

#Plotting parameters: 

x <- seq(1200,1900,10)
y1ver <- logverlower(x)
y2ver <- logverupper(x)

y1know <-logsaberlower(x)
y2know <- logsaberupper(x)

# Plot:

mood.verbs.factive <- ggplot()+  
              geom_point(data=plot.data.verbs.factive,
                         aes(Period, mean,size = n, color=Mverbl))+scale_size_area(max_size=13)+ 
              labs(title="Proportion Mood: I don't V that p + Fitted logistic(1200-1900)", 
                    x = " ", 
                    y="Mood Proportion (1=ind,0=subj)")+
             scale_color_manual(values=c("#E66100", "#5D3A9B" ))+
             geom_function(fun=logsaber, color = "orange")+
              geom_function(fun=logver, color = "purple")+
            ylim(0,1)+
            labs(size="size", colour="Verb")+
            geom_segment(aes(x = 1428, y = 0.50, xend = 1520, yend = 0.50))+
            geom_segment(aes(x=1428, y=0.45, xend = 1428, yend=0.55))+
            geom_segment(aes(x=1520, y=0.45, xend = 1520, yend=0.55))+
            geom_segment(aes(x = 1520, y = 0.50, xend = 1539, yend = 0.50), linetype="dotted")+
            geom_segment(aes(x=1539, y=0.45, xend = 1539, yend=0.55),linetype="dotted")+
            geom_polygon(aes(c(x,rev(x)),c(y2ver,rev(y1ver))),fill="purple",alpha=0.1)+
            geom_polygon(aes(c(x,rev(x)),c(y2know,rev(y1know))),fill="orange",alpha=0.1)
mood.verbs.factive
```


#### Second Attempt: fitting only the drop 

```{r}
#Fit the data from 1400 onwards only:

clean.data.know.1400 <- subset(clean.data, 
                    Mverbl=="saber (know)"& # Only know
                    Construction=="1st present (FP)"& # only first person
                      Period!="1200"& # data from 1400 onwards
                    Period!="1300"
                   )

modelknow1400 <- nls(Emood2 ~ logarithm(Period,k,s), data=clean.data.know.1400, start=list(s=0.024, k=1439))
summary(modelknow1400)

clean.data.see.1500<- subset(clean.data, 
                    Mverbl=="ver (see)"& # Only the verb see
                    Construction=="1st present (FP)"& # only first person constructions
                    Period!="1200"& # data from 1500 onwards.
                    Period!="1300"&
                    Period!="1400"
                   )

modelsee <- nls(Emood2 ~ logarithm(Period,k,s), data=clean.data.see.1500, start=list(s=0.008, k=1505))
summary(modelsee)
```

Plotting the results

```{r}
logsaber1500 <-function(x) {1/(1+exp(0.023*(x-1438)))}
logsaber1500upper <-function(x) {1/(1+exp(0.027*(x-1447)))}
logsaber1500lower <-function(x) {1/(1+exp(0.019*(x-1429)))}

logver1500 <-function(x) {1/(1+exp(0.012*(x-1539)))}
logver1500lower <-function(x) {1/(1+exp(0.009*(x-1523)))}
logver1500upper <-function(x) {1/(1+exp(0.015*(x-1555)))}

#Calculating the error upper bouns: 0.019=> 92 years(1429+117=1546), 0.023=>76 years(1438+76=1514); 

x <- seq(1200,1900,10)
y1ver1500 <- logver1500lower(x)
y2ver1500 <- logver1500upper(x)

y1know1400 <-logsaber1500lower(x)
y2know1400 <- logsaber1500upper(x)

#delete points not used
plot.data.verbs.factive2 <- subset(plot.data.verbs.factive, Period!="1200" &
                                     Period!="1300"
                                   )
plot.data.verbs.factive3 <- filter(plot.data.verbs.factive2, (Period!="1400"| Mverbl=="saber (know)"))

# Plotting the graph:

moodverbs.factive1400 <- ggplot()+  
              geom_point(data=plot.data.verbs.factive3,
                         aes(Period, mean,size = n, color=Mverbl))+scale_size_area(max_size=13)+ 
              labs(title="Proportion Mood: I don't V that p + Fitted logistic (1400/1500-1900)", 
                    x = " ", 
                    y="Mood Proportion (1=ind,0=subj)")+
             scale_color_manual(values=c("#E66100", "#5D3A9B" ))+
              geom_function(fun=logsaber1500, color = "orange")+
              #geom_function(fun=logsaber1500upper, color = "orange", linetype="dotted")+
              #geom_function(fun=logsaber1500lower, color = "orange", linetype="dotted")+
              geom_function(fun=logver1500, color = "purple")+
              #geom_function(fun=logver1500lower, color = "purple", linetype="dotted")+
              #geom_function(fun=logver1500upper, color = "purple", linetype="dotted")+
              #stat_function(fun=logver1500upper, geom="area", fill="#84CA72", alpha=0.2)+
            ylim(0,1)+
            labs(size="size", colour="Verb")+
            geom_segment(aes(x = 1438, y = 0.50, xend = 1514, yend = 0.50))+
            geom_segment(aes(x=1438, y=0.45, xend = 1438, yend=0.55))+
            geom_segment(aes(x=1514, y=0.45, xend = 1514, yend=0.55))+
            geom_segment(aes(x = 1514, y = 0.50, xend = 1546, yend = 0.50),color="black",linetype="dotted")+
            geom_segment(aes(x=1546, y=0.45, xend = 1546, yend=0.55),linetype="dotted")+
            geom_polygon(aes(c(x,rev(x)),c(y2ver1500,rev(y1ver1500))),fill="purple",alpha=0.1)+
  geom_polygon(aes(c(x,rev(x)),c(y2know1400,rev(y1know1400))),fill="orange",alpha=0.1)
moodverbs.factive1400

```
#### Combining the two attempts

```{r}
combined.models <- ggarrange(mood.verbs.factive,moodverbs.factive1400)
```

```{r}
ggsave(combined.models, file="fitted-loss-1stperson2.png",width = 13,height = 4)
```

## Appendix I: -ra subjunctive forms

```{r}
ex.datarase <- subset(data, Mverbtype =="ind" & 
                     Emood !="Na" & 
                     Emood !="NA" &
                     Emood !="inf" &
                     Emoodrase !="ra" & # excludes the -ra forms 
                     MClauseType=="noninterrogative" &
                     EClauseType=="unambiguous"
                   )

# Changing the indicative to 1 and subjunctive to 0:
ex.datarase$Emood2 <- ifelse(ex.datarase$Emood == "subj", 0, 1)

#changing name factive and non-factive
ex.datarase$Mverbclass2 <- ifelse(ex.datarase$Mverbclass == "factive", "semi-factive", "non-factive")
ex.datarase$Construction <- ifelse(ex.datarase$Construction == "1st present", "1st present (FP): without embedded -ra", "other (O): without embedded -ra")

# Converting the into a numeric value:
ex.datarase$Emood2 <- as.numeric(as.character(ex.datarase$Emood2))
ex.datarase$Period <- as.factor(as.character(ex.datarase$Period))

# new dataframe with the means of embedded mood per period, verb and type of matrix subject(persona)
plot.datafactiverase <- ddply(ex.datarase, .(Period, Construction,Mverbclass2
),summarize, mean = mean(as.numeric(as.character(Emood2)), na.rm = T), n = sum(!is.na(as.numeric(as.character(Emood2)))))


# original data: 
clean.data$Mverbclass2 <- ifelse(clean.data$Mverbclass == "factive", "semi-factive", "non-factive")
plot.datafactive <- ddply(clean.data, .(Period, Construction,Mverbclass2),
                          summarize, 
                          mean = mean(as.numeric(as.character(Emood2)), na.rm = T),
                          n = sum(!is.na(as.numeric(as.character(Emood2))))
                          )
plot.datafactive$Period <- as.factor(plot.datafactive$Period)


moodfactiverase <- ggplot()+  
  geom_point(data=plot.datafactiverase, aes(Period, mean, color=Construction), alpha=0.6, size=4)+
  geom_point(data=plot.datafactive, aes(Period, mean, color=Construction), alpha=0.6, size=4)+
  facet_wrap(~factor(Mverbclass2, c("semi-factive","non-factive")), ncol=2)+scale_size_area(max_size=13)+ 
  labs(title="Proportion Mood (with/without -ra forms): not V(ind) that p(ind/subj).", 
       x = " ", 
       y="Mood Proportion (1=ind,0=subj)")+
  scale_color_manual(values=c("black","gray", "#5F9EA0", "#3589de"))+
  ylim(0,1)+
  labs(colour="Construction")
  

moodfactiverase

```
```{r}
ggsave(moodfactiverase, file="diachrony-verb-classes-rase2.png", width = 11, height = 4)
```

## Appendix II: ambiguous sentences
```{r}
ex.data.ambiguous <- subset(data, Mverbtype =="ind" & 
                     Emood !="Na" & 
                     Emood !="NA" &
                     Emood !="inf" &
                     MClauseType=="noninterrogative" 
                   )

# Changing the indicative to 1 and subjunctive to 0:
ex.data.ambiguous$Emood2 <- ifelse(ex.data.ambiguous$Emood == "subj", 0, 1)

#changing name factive and non-factive
ex.data.ambiguous$Mverbclass2 <- ifelse(ex.data.ambiguous$Mverbclass == "factive", "semi-factive", "non-factive")
ex.data.ambiguous$Construction <- ifelse(ex.data.ambiguous$Construction == "1st present", "1st present (FP): with ambiguous sentences", "other (O): with ambiguous sentences")

# Converting the into a numeric value:
ex.data.ambiguous$Emood2 <- as.numeric(as.character(ex.data.ambiguous$Emood2))
ex.data.ambiguous$Period <- as.factor(as.character(ex.data.ambiguous$Period))

# new dataframe with the means of embedded mood per period, verb and type of matrix subject(persona)
plot.datafactive.ambiguous <- ddply(ex.data.ambiguous, .(Period, Construction,Mverbclass2
),summarize, mean = mean(as.numeric(as.character(Emood2)), na.rm = T), n = sum(!is.na(as.numeric(as.character(Emood2)))))

```


```{r}
moodfactiveambiguous <- ggplot()+  
  geom_point(data=plot.datafactive.ambiguous, aes(Period, mean, color=Construction), alpha=0.6, size=4)+
  geom_point(data=plot.datafactive, aes(Period, mean, color=Construction), alpha=0.6, size=4)+
  facet_wrap(~factor(Mverbclass2, c("semi-factive","non-factive")), ncol=2)+scale_size_area(max_size=13)+ 
  labs(title="Proportion Mood (with/without ambiguous sentences): not V(ind) that p(ind/subj).", 
       x = " ", 
       y="Mood Proportion (1=ind,0=subj)")+
  scale_color_manual(values=c("black","gray", "#5F9EA0", "#3589de"))+
  ylim(0,1)+
  labs(colour="Construction")
  
moodfactiveambiguous
```
```{r}
ggsave(moodfactiveambiguous, file="diachrony-verb-classes-ambiguous2.png", width = 11, height = 4)
```

















