---
title: "The Development of Polarity Subjunctive"
author: "Raquel Montero Estebaranz"
date: "03.04.2024"
output: 
  html_document: 
      toc: true
      toc_float: true
      theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center", warning = FALSE, message = FALSE)

```
```{css, echo=FALSE}
h1,h4 {
  text-align: center;
}
p {
  text-align: justify; 
}
```

## 1. Introduction

This file shows the code used to analyse the data of the verb "say" for Chapter 5 of the thesis entitled "Mood alternations: a synchronic and diachronic study of negated complement clauses". The data as well as the annotation guidelines can be found at: https://github.com/Raquel-Montero


The following are the Packages that will be used:
```{r}
library(readr)
library(carData)      # for cat package
library(car)          # Anova function
library(dplyr)        # Operations
library(plyr)         # for ddply 
library(ggplot2)      # to use ggplot
library(sjPlot)       # to change the font
library(Matrix)       # for lme4 package
library(lme4)         # to calculate lmer models
library(lmerTest)     # for the p values
library(mgcv)         # for Gam model
library(itsadug)      # for Gam model
library(tidygam)      # for Gam plotting
library(tidymv)       # For Gam plotting: https://www.rdocumentation.org/packages/tidymv/versions/3.4.2
library(ggpubr)
library(LaplacesDemon) #for the function invlogit
```


## 2. Data

Loading the Data: 

```{r}
data <- read_csv("DiachronyAllVerbs.csv", 
                  show_col_types = FALSE,
                  locale = locale(encoding =  "ISO-8859-1"))

```



# 3. The verb say

```{r}
ex.data <- subset(data, Mverbtype =="ind" & 
                     Mverbclass!="Na"& # not one of the four verbs
                     Emood !="Na" & 
                     Emood !="NA" &
                     Emood !="inf" &
                     MClauseType=="noninterrogative" &
                     EClauseType=="unambiguous"&
                     Mverbl=="decir (say)"
                   )

# Changing the indicative to 1 and subjunctive to 0:
ex.data$Emood2 <- ifelse(ex.data$Emood == "subj", 0, 1)
# Converting the into a numeric value:
ex.data$Emood2 <- as.numeric(as.character(ex.data$Emood2))

#changing names of values so that they are better for plotting: 
ex.data$Mverbtype2 <- ifelse(ex.data$Mverbclass == "factive", "Semi-factive", "Non-factive)")
ex.data$Construction <- ifelse(ex.data$Construction == "1st present", "1st present (FP)", "other (O)")

```

Data for plotting: 

```{r}
# new dataframe with the means of embedded mood per period, verb class/verb and construction:

plot.data.say <- ddply(ex.data, .(Period, Construction),
                          summarize, 
                          mean = mean(as.numeric(as.character(Emood2)), na.rm = T),
                          n = sum(!is.na(as.numeric(as.character(Emood2))))
                          )

```

Plotting the results:

```{r}
#Font for the graph:
windowsFonts("Garamond" = windowsFont("Garamond"))

set_theme( base = theme_bw(base_family = 'Garamond'),
  title.size = 1.4,
  axis.title.size = 1.4,
  axis.textsize = 1.1,
  legend.size = 1,
  legend.title.size = 1,
  geom.label.size = 3
)

# Plot 1:
plot.data.say$Period <- as.factor(plot.data.say$Period) # Period as factor

plot.say <- ggplot()+  
                geom_point(data=plot.data.say, 
                           aes(Period, mean, size = n, color=Construction))+ #main data 
                scale_size_area(max_size=20,limits=c(1,1100))+
                labs(title="Proportion Mood: not say(ind) that p(ind/subj)",  # axis
                        x = "", 
                       y="Proportion embedded mood (1=ind, 0=subj)")+
                scale_color_manual(values=c("black", "#4c8c44d9" ))+ # colors
                labs(size="size", colour="Construction")+ # labels legends
                ylim(0,1)
  
plot.say


```
```{r}
#Saving the plot:
ggsave(plot.say, file="plot-say.png", width = 6.5, height= 4)
```

## 3.1. Speed of change: logistic
```{r}

# Data for Construction O:
ex.dataB <- subset(ex.data, Construction=="other (O)")

# defining a logarithmic function:
logarithmB <- function(Period,k,s,m)(1/((1/m)+exp((Period-k)/s)))

# Fiting the model to the data:
#Initial values: m = 0.79; s = 163; k = 1980 (this values were obtained by using the graphic calculator Desmos)
# calculating the vales with nls:
modelB <- nls(Emood2 ~ logarithmB(Period,k,s,m), data=ex.dataB, start=list(s=163, k=1980, m=0.79))
summary(modelB) 

#Obtained values --> 
                    #K = 1946+-51*2= 102; 
                    #s = 164 +- 90*2=180 --> [0-344];
                    #m = 0.82 +-0.07*2 = 14 

# Some procedure for Construction FP:

ex.dataA <- subset(ex.data, Construction=="1st present (FP)")

logarithmA <- function(Period,k,s,m)(1/((1/m)+exp((Period-k)/s)))

modelA <- nls(Emood2 ~ logarithmA(Period,k,s,m), data=ex.dataA, start=list(s=88, k=1696, m=0.51))
summary(modelA)

## k =1700 +-48*2= 96=[],
#s = 91+- 35*2 =70, 
# m = 0.56 +-0.06*2 =0.12

```

Ploting the model fit to the data:
```{r}
#Main functions:
fitB <- function(x){(1/((1/0.82)+exp((x-1946)/164)))}
fitA <- function(x) {1/((1/0.56)+exp((x-1700)/91))}

#Confidence intervals for the speed of Construction O:

fitBupper <-function(x){(1/((1/0.96)+exp((x-2048)/344)))}
fitBlower <-function(x){(1/((1/0.68)+exp((x-1844)/-20)))}
#Confidence intervals for the speed of construction FP:
fitAupper <- function(x) {1/((1/0.68)+exp((x-1796)/161))}
fitAlower <- function(x) {1/((1/0.44)+exp((x-1604)/21))}

#equations for plotting the confiddence intervals:
x <- seq(1200,1900,10)
y1O <- fitBupper(x)
y2O <- fitBlower(x)

x <- seq(1200,1900,10)
y1FP <- fitAupper(x)
y2FP <- fitAlower(x)

plot.data.say$Period <- as.numeric(as.character(plot.data.say$Period))


plot.say.fit <- ggplot()+  
                geom_point(data=plot.data.say, 
                           aes(Period, mean, size = n, color=Construction))+#main data 
                scale_size_area(max_size=20,limits=c(1,1100))+  # controls the maximum size of the points.
                geom_function(fun=fitB, color = "#4c8c44d9")+
                geom_function(fun=fitA, color = "black")+
                labs(title="Proportion Mood: not say(ind) that p(ind/subj)",  # axis
                      x = " ", 
                       y="Proportion embedded mood (1=ind,0=subj)")+
                scale_color_manual(values=c("black", "#4c8c44d9" ))+ # colors
                labs(size="size", colour="Construction")+ # labels legends
                ylim(0,1)+
                geom_polygon(aes(c(x,rev(x)),c(y2O,rev(y1O))),fill="#4c8c44d9",alpha=0.1)+
                geom_polygon(aes(c(x,rev(x)),c(y2FP,rev(y1FP))),fill="gray",alpha=0.5)
                #geom_polygon(aes(c(x,rev(x)),c(y2decir,rev(y1decir))),fill="black",alpha=0.1)
                
  
plot.say.fit


```
```{r}
ggsave(plot.say.fit, file="say-fit-CI.png", width = 8, height = 5)
```
The confidence interval of fitting the logistic to Construction A show that there might be something wrong with the model.

## 3.2. The speed of change: Laplace
```{r}
#Define a function to fit data to: 
laplace <- function(Period,m,k,s)(m*exp(-(abs(Period-k)/s)))
                                  
                                  
#NLLS: 
modellaplaceB <- nls(Emood2 ~ laplace(Period,m,k,s), data=ex.dataB, start=list(m=0.43,k=1572, s=200))
summary(modellaplaceB) 
# m= 0.90 +-(0.04*2)0.08
# k = 1553 +-(17*2)34
# s= 637 +-(116*2)232


modellaplaceA <- nls(Emood2 ~ laplace(Period,m,k,s), data=ex.dataA, start=list(m=0.43,k=1572, s=200))
summary(modellaplaceA) 
# speed of change: 325 +-160(80x2) --> 485
#m= 0.60 +-(0.06*2)0.12
#k= 1512 +-(24*2)48
#s= 325+- (83*2)166
# The speeds of change given by the model also overlap they are not statistically significant from one another.


fitBLaplace <- function(x){0.90* exp(-(abs(x-1553)/637))}
fitALaplace <- function(x){0.60* exp(-(abs(x-1512)/325))}




plot.say.fit.laplace <- ggplot()+  
                geom_point(data=plot.data.say, 
                           aes(Period, mean, size = n, color=Construction))+#main data 
                scale_size_area(max_size=20,limits=c(1,1100))+  # controls the maximum side of the points.
                geom_function(fun=fitBLaplace, color = "#4c8c44d9")+
                geom_function(fun=fitALaplace, color = "black")+
                labs(title="Proportion Mood: not say(ind) that p(ind/subj)",  # axis
                      x = " ", 
                       y="Proportion embedded mood (1=ind,0=subj)")+
                scale_color_manual(values=c("black", "#4c8c44d9" ))+ # colors
                labs(size="size", colour="Construction")+ # labels legends
                ylim(0,1)
                #geom_polygon(aes(c(x,rev(x)),c(y2decir,rev(y1decir))),fill="black",alpha=0.1)
                
  
plot.say.fit.laplace

```
Laplace with confidence intervals:
```{r}

#Main functions:
fitBLaplace <- function(x){0.90* exp(-(abs(x-1553)/637))}
fitALaplace <- function(x){0.60* exp(-(abs(x-1512)/325))}

#Confidence intervals for the speed of Construction O:
fitBLaplaceupper <-function(x){0.98* exp(-(abs(x-1567)/869))}
fitBLaplacelower <-function(x){0.82* exp(-(abs(x-1519)/405))}
#Confidence intervals for the speed of construction FP:
fitALaplaceupper <- function(x){0.72* exp(-(abs(x-1560)/491))}
fitALaplacelower <- function(x){0.48* exp(-(abs(x-1464)/159))}

#equations for plotting the confiddence intervals:
x <- seq(1200,1900,10)
y1Olaplace <- fitBLaplaceupper(x)
y2Olaplace <- fitBLaplacelower(x)

x <- seq(1200,1900,10)
y1FPlaplace <- fitALaplaceupper(x)
y2FPlaplace <- fitALaplacelower(x)

#plot.data.say$Period <- as.numeric(as.character(plot.data.say$Period))
plot.say.fit.laplace.ci <- ggplot()+  
                geom_point(data=plot.data.say, 
                           aes(Period, mean, size = n, color=Construction))+#main data 
                scale_size_area(max_size=20,limits=c(1,1100))+  # controls the maximum side of the points.
                geom_function(fun=fitBLaplace, color = "#4c8c44d9")+
                geom_function(fun=fitALaplace, color = "black")+
                labs(title="Proportion Mood: not say(ind) that p(ind/subj)",  # axis
                      x = " ", 
                       y="Proportion embedded mood (1=ind,0=subj)")+
                scale_color_manual(values=c("black", "#4c8c44d9" ))+ # colors
                labs(size="size", colour="Construction")+ # labels legends
                ylim(0,1)+
              geom_polygon(aes(c(x,rev(x)),c(y2Olaplace,rev(y1Olaplace))),fill="#4c8c44d9",alpha=0.1)+
  geom_polygon(aes(c(x,rev(x)),c(y2FPlaplace,rev(y1FPlaplace))),fill="gray",alpha=0.5)
                #geom_polygon(aes(c(x,rev(x)),c(y2decir,rev(y1decir))),fill="black",alpha=0.1)
                
  
plot.say.fit.laplace.ci

```


The confidence intervals are better in the case showing that there the model is behaving as expected.

```{r}
ggsave(plot.say.fit.laplace, file="say-fit-laplace.png", width = 8, height = 5)
ggsave(plot.say.fit.laplace.ci, file="say-fit-laplace-ci.png", width = 8, height = 5)
```


## 3.3. Comparing the AIC of the models.
```{r}
#AIC: 
AIC(modelB,modellaplaceB)#Laplace is a better model with AIC 645 than logistic with 654 
AIC(modelA,modellaplaceA)# Laplace is a worse model with AIC 550 vs logistic 541

```


## 3.4. The speed of change:  Hubbert Curve
```{r}
#Define a function to fit data to: 
hubbert <- function(Period,k,m,j,r){(exp(-(Period-k)/(m)))/(1+exp(-(Period-j)/(r)))^2}
                                  
                                  
#NLLS: 
modelhubbertB <- nls(Emood2 ~ hubbert(Period,k,m,j,r), data=ex.dataB, start=list(k=1349,j=1465, m=-140, r=-100))
summary(modelhubbertB)
# Values of the model: 
# K 1144 +- 123*2
# j 1335 +- 411*2
# m -178 +- 192*2  slope for the first half of the curve (there is too little data)
# r -237 +- 108*2 

modelhubbertA <- nls(Emood2 ~ hubbert(Period,k,m,j,r), data=ex.dataA, start=list(k=1349,j=1465, m=-140, r=-140))
summary(modelhubbertA)
# k 1506 +- 563*2
# j 1666 +- 410*2
# m -361 +- 766*2 --> in the early period there is too little data 
# r -158 +- 53*2

```
Plotting the curves:

```{r}
#Plotting the curves: 

#Main functions:
fitBhubbert <- function(x){(exp(-(x-1144)/(-178)))/(1+exp(-(x-1335)/(-237)))^2}
fitAhubbert <- function(x){(exp(-(x-1505)/(-360)))/(1+exp(-(x-1666)/(-158)))^2}


plot.say.fit.hubbert <- ggplot()+  
                geom_point(data=plot.data.say, 
                           aes(Period, mean, size = n, color=Construction))+#main data 
                scale_size_area(max_size=20,limits=c(1,1100))+  # controls the maximum side of the points.
                geom_function(fun=fitBhubbert, color = "#4c8c44d9")+
                geom_function(fun=fitAhubbert, color = "black")+
                labs(title="Proportion Mood: not say(ind) that p(ind/subj)",  # axis
                      x = " ", 
                       y="Proportion embedded mood (1=ind,0=subj)")+
                scale_color_manual(values=c("black", "#4c8c44d9" ))+ # colors
                labs(size="size", colour="Construction")+ # labels legends
                ylim(0,1)
                
  
plot.say.fit.hubbert


```

```{r}
ggsave(plot.say.fit.hubbert, file="say-fit-hubbert.png", width = 8, height = 5)
```


## 3.5. Comparing the models
```{r}
AIC(modelA,modelhubbertA) # no difference in the models: 541 vs 542 less than 2 point difference
AIC(modelB, modelhubbertB) # hubbert model is preffered
```









